Spatial variation in predator diel activity patterns -- feral cats avoid red foxes in time, not space

### 2022-04-14 manuscript/main_text.Rmd
post-phd-getting-ready-for-submission

---
title: "Spatial variation in predator diel activity patterns -- feral cats avoid red foxes in time, not space"
author:
  - name: Matthew W. Rees
    email: matt.wayne.rees@gmail.com
    affiliation: UOM
    footnote: 1
  - name: Jack H. Pascoe
    email: 
    affiliation: CEC
  - name: Mark Le Pla
    email: 
    affiliation: CEC
  - name: Alan Robley
    email: 
    affiliation: ARI
  - name: Emma K. Birnbaum
    email: 
    affiliation: CEC
  - name: Brendan A. Wintle
    email: 
    affiliation: UOM
  - name: Bronwyn A. Hradsky
    email: 
    affiliation: UOM
    footnote: 
address:
  - code: UOM
    address: Quantitative & Applied Ecology Group, School of Ecosystem and Forest Science, The University of Melbourne, Parkville, VIC, Australia
  - code: CEC 
    address: Conservation Ecology Centre, Otway Lighthouse Rd, Cape Otway, VIC, Australia
  - code: ARI 
    address: Department of Environment, Land, Water and Planning, Arthur Rylah Institute for Environmental Research, Heidelberg, Australia
footnote:
  - code: 1
    text: "Corresponding Author"
abstract: 
keywords: "feral cat; diel activity patterns; generalised additive model; invasive predator;  intraguild predator interactions; mesopredator release; red fox; lethal predator control"
journal: "?"
date: "`r Sys.Date()`"
bibliography: /Users/mrees2/Dropbox/personal/matt/github/phd-thesis/index/bib/thesis.bib
header-includes:
    - \usepackage{setspace}\doublespacing
    - \usepackage{float}
    - \floatplacement{figure}{H}
    - \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}} \setcounter{section}{0} \renewcommand{\thesection}{S\arabic{section}}}
    - \usepackage{lineno}
linenumbers: false
csl: journal-of-applied-ecology.csl
geometry: margin=1.1in
fontsize: 12pt
output:
  bookdown::pdf_book:
    base_format: rticles::elsevier_article
editor_options: 
  chunk_output_type: inline
---

@@

## Study area and camera-trapping 
We compiled data from multiple smaller-scale camera-trap studies across two regions in south-west Victoria, Australia: the Glenelg region and Otway Ranges (Fig. \@ref(fig:diel-map)). Introduced foxes and cats are the only medium-large functional mammalian terrestrial carnivores here: native dingoes *Canis familiaris* are long-absent throughout, while tiger quolls *Dasyurus maculatus* are long-absent in the Glenelg region and likely functionally extinct in the Otway Ranges (last confirmed sighting in 2014). In broad sections of each region, government land managers conduct ongoing targeted lethal fox control for biodiversity conservation. Poison-baits containing 3 mg of sodium mono-fluroacetate ('1080') are buried at a depth of 12 - 15 cm at 1-km intervals along accessible forest tracks and roads. Different road densities result in variable densities of poison-baits. Managers also frequently implement prescribed fire across both regions, primarily to reduce fuel loads to prevent large wildfires.

### Glenelg region
In the Glenelg region, large patches of natural vegetation are fragmented by pastoral farming and residential properties (Fig. \@ref(fig:diel-map)). Foxes in three distinct forest blocks in this region have been subject to poison-baiting since October 2005, with fortnightly bait replacements [@robley2014long]. These forest blocks, along with three similar, unbaited forest blocks to the north have been simultaneously surveyed annually under the 'Glenelg Ark' fox control program since 2005 [40 sites per block; @robley2020glenelg]. Hair-tubes were used to monitor species from 2005 - 2013 [presented in @robley2014long], replaced by camera-traps from 2013; here we present camera-trap data from 2013 - 2019 [@robley2020glenelg]. We also included a further 425 camera-trap deployments at unique locations from early 2018 (M.W.R., PhD surveys). This totals 2039 camera-trap deployments in the Glenelg region, collected in a control-impact experimental design (foxes had been continuously controlled for at 8 - 14 years in the treatment landscapes at the time of these surveys).

@@

#### Camera-trap set-ups
All camera-trap deployments consisted of a Reconyx (Holmen, Wisconsin) brand camera-trap (white or infrared flash), attached to a tree or a metal picket, facing a lure. The Glenelg Ark and Otway Ark fox monitoring programs positioned camera-traps at least 40 cm above ground on a tree or a metal picket and angled downwards toward a lure approximately 1 - 1.5 m away [@robley2020glenelg; @robley2019otway]. The lures consisted of peanut butter, golden syrup and rolled oats mixed into a small ball, placed within a tea strainer or PVC pipe container and secured either to the ground, or 20 - 60 cm above ground on a wooden stake. The M.W.R. PhD surveys across both regions positioned camera-traps lower on a tree (around 15 - 30 cm above the ground) angled only slightly downwards toward a tuna oil lure approximately 2 - 2.5 m away [detailed in @rees2019unexpectedly]. Camera-traps were active for an average of 47 days (maximum 93 days), totalling 172,052 trap-nights.

@@

#### How does predator overall activity and diel activity patterns vary across vegetation types? (model 2)
Predator activity varied across space; we hypothesised that this was partly due to differences in vegetation type, based both on the observed spatial patterns and because vegetation type is a major driver of understorey habitat structure and prey occurrence in these regions [@swan2015predicting; @hradsky2017bayesian]. To test whether the diel activity pattern of each predator varied among vegetation types, we identified the Ecological Vegetation Class group [hereafter 'vegetation type'; standard units for vegetation classification in Victoria; @delwp2020bioregions] for each unique camera-trap site, totalling eight vegetation types. As rainforests are interspersed (primarily in low lying gullies) at fine-scales throughout wet and damp forests in the south-eastern Otway Ranges, we merged them together (hereafter referred to as 'wet forests'). We then estimated predator activity across vegetation types using a hierarchical model specification: a global smoother for hour (i.e., average response) and group-level smoothers with shared wiggliness for the seven vegetation types ['model GS' detailed in @pedersen2019hierarchical]. We also included a random effect to account for differences in overall activity levels between the two regions.

@@

RESULTS
============

Overall, we collated 5449 and 2202 independent detections of foxes and cats, respectively (separated by at least 30 minutes) from 172,052 camera-trap nights (Table \@ref(tab:diel-tab1)).

## How does predator overall activity and diel activity patterns vary across space? (model 1)
Predator activity varied considerably across space and throughout the 24-hour daily cycle, and there was some variation in the predator diel activity patterns across space. On average, both predators showed similar diel activity patterns: mostly nocturnal with peaks in activity around sunrise and sunset (i.e., crepuscular; Fig. \@ref(fig:diel-veg)i). The main difference between the species was that fox activity peaked just after sunset and they were less likely to be active during the day than cats. Cats also tended to be more active at sunset relative to sunrise.

@@

Variation in diel activity patterns across space, as well as differences in overall activity between the predators, was strongest in the Otway Ranges. For example, overall fox activity (Fig. \@ref(fig:diel-space-o-fox)) and diel activity pattern strength(Fig. \@ref(fig:diel-space)b) were lowest in the south-west Otway Ranges, while feral cat overall activity (Fig. \@ref(fig:diel-space-o-cat)) and diel activity pattern strength (Fig. \@ref(fig:diel-space)b) were highest in that subregion.

## How does predator overall activity and diel activity patterns vary across vegetation types? (model 2)

@@

Diel activity patterns for foxes were similar across all vegetation type except wet forests; in wet forests, foxes were consistently active throughout the daily cycle (Fig. \@ref(fig:diel-veg)a). On the other hand, cats were nocturnal (and most active) in wet forests, but largely crepuscular in all other vegetation types (Fig. \@ref(fig:diel-veg)b). For both predators, the random effect for region (Glenelg or Otways) in the vegetation models shrank to near-zero=, indicating all variation between the regions was explained by the vegetation covariate and site random intercept.

## Do feral cats avoid foxes in space or time? (model 3)
Cat spatial activity was relatively unaffected by the fox activity in both habitat types of the Otway Ranges; and, if anything, increased with increasing adjusted fox counts in the Glenelg region (Fig. \@ref(fig:diel-cat-fox)), indicating cats did not avoid foxes spatially.

@@

A key question in ecological theory is whether animals are evolutionary hardwired to occupy particular temporal niches, or have circadian rhythms that are responsive to changing environmental conditions and interactions with other species [@daan1981adaptive; @schoener1974compression; @lima1990behavioral]. Here we demonstrate that diel activity patterns are not fixed, but vary across space based on landscape context and fear. In our study, sympatric invasive predators had similar diel activity patterns when averaged across broad regions [i.e., high circular overlap; Fig. \@ref(fig:diel-space-time-marginal); as did @roshier2021space], but behaviours varied considerably within landscapes. Fox daily activity patterns were most strongly tied to the daily cycle in dry habitat types, but showed little diel activity pattern in the wet forests (Fig. \@ref(fig:diel-veg)a). In contrast, cats were mostly nocturnal in wet forests but crepuscular in dry vegetation types (Fig. \@ref(fig:diel-veg)b). Within broad habitat types, cats altered their diel activity patterns at sites with higher fox activity (Fig. \@ref(fig:diel-cat-fox)). Control programs that reducing invasive fox activity are therefore likely to change cat diel activity patterns, which in turn may alter cat impacts on native prey species. Quantifying changes in diel activity patterns provides important context for understanding species interactions, which is key for effective ecosystem management [@gaynor2021applied].

Shifting diel activity patterns may facilitate spatial coexistence of dominant and subordinate species [@carothers1984time]. For cats, altering diel activity patterns to less-preferred times of day may be worthwhile to persist in high-quality habitat. Few studies have demonstrated predator-induced shifts in diel activity [@kronfeld2003partitioning], but notably, ship rats *Rattus norvegicus* were also found to switch from nocturnal to diurnal behaviour in response to fox activity [@fenn1995use]. For cats in our study, a switch to diurnal behaviour where fox activity was high in the dry vegetation types may have been facilitated by the higher abundance of reptiles in these habitat types relative to wet forests [which are mostly diurnal; @woinarski2018reptiles]. At wet forest sites with high fox activity, cats concentrated their activity away from sunrise and sunset towards midnight, despite a diurnal shift appearing to similarly reduce the risk of a fox encounter. In this situation, we expect becoming more nocturnal to be favourable over a diurnal shift because this is when small mammals are active and cats would be least visible to foxes [cats here mostly had black or grey coats]. Overall fox activity was twice as low in the wet forests relative to dry habitat types, and so cats were likely under less pressure to radically alter their diel activity patterns. Understanding how these potential avoidance behaviours impacts native prey is a key research priority to improve invasive predator management.

Cats may have avoided foxes in time, but we saw no sign of spatial avoidance (i.e. no fine-scale negative association between fox and cat overall activity (Fig. \@ref(fig:diel-cat-fox)b:c). Overlap in spatial activity at fine scales (i.e., at the camera-trap sites) may have been facilitated by temporal avoidance [@kronfeld2003partitioning]. However, we considered spatial avoidance across the entire survey duration (averaging 47 days). Cats may indeed avoid foxes spatially, but transiently on the scale of hours to days—after all, how is a cat to know where to avoid a fox without encountering signs of one? Short-term spatial avoidance is quite plausible given foxes mark territories using scats and odours, which cats could tangibly associate with high risk shortly after. Temporary spatial avoidance could be tested using decay curves [e.g., @niedballa2019assessing]. Alternatively, no sign of spatial avoidance between these invasive predators could also be an artefact of the quality of camera set-up and hence detectability. The 3667 camera-traps were deployed by numerous people, and the quality of set-ups differed considerably in terms of detecting predators. Camera-traps that angled only slightly downwards, rather than upwards or strongly downwards, seemed most effective at detecting both predator species (M.W Rees, personal observation). Predator interactions are routinely inferred through spatial associations between species; however such analyses are subject to numerous pitfalls which can make inference unreliable [reviewed in @guillaume2020co].

@@

Animal diel activity patterns can be complex, varying across space, habitat types and threat-levels. Despite telling an important story about how animals interact with each other and the environment, detection times are commonly discarded from statistical analyses of camera-trap data. In the rare instances that they are considered, diel activity patterns are predominantly estimated at the population-level, overlooking finer-scale behaviours that can affect fitness, survival and ecosystem-impacts. Our results demonstrate the importance of (a) considering diel activity in regards to species interactions, (b) modelling *changes* in animal behaviour rather than overlap with other species, and (c) testing avoidance behaviours within a joint spatiotemporal framework. Our study adds to the limited body of evidence that top predators can produce a landscape of fear which is powerful enough to reverse the diel activity patterns of subordinate species [@kronfeld2003partitioning].

@@

@@

# load records
fits_summaries <- read_csv("../derived_data/model_summaries.csv") %>%
  dplyr::select(-c(X1))

>>>>

---
title: "Spatial variation in predator diel activity patterns -- feral cats avoid red foxes in time, not space"
author:
  - name: Matthew W. Rees
    email: matt.wayne.rees@gmail.com
    affiliation: UOM
    footnote: 1
  - name: Brendan A. Wintle
    email: 
    affiliation: UOM
  - name: Jack H. Pascoe
    email: 
    affiliation: CEC
  - name: Mark Le Pla
    email: 
    affiliation: CEC
  - name: Emma K. Birnbaum
    email: 
    affiliation: CEC
  - name: Bronwyn A. Hradsky
    email: 
    affiliation: UOM
    footnote: 
address:
  - code: UOM
    address: Quantitative & Applied Ecology Group, School of Ecosystem and Forest Science, The University of Melbourne, Parkville, VIC, Australia
  - code: CEC 
    address: Conservation Ecology Centre, Otway Lighthouse Rd, Cape Otway, VIC, Australia
footnote:
  - code: 1
    text: "Corresponding Author"
abstract: 
keywords: "feral cat; diel activity patterns; generalised additive model; invasive predator;  intraguild predator interactions; mesopredator release; red fox; lethal predator control"
journal: "?"
date: "`r Sys.Date()`"
bibliography: /Users/mrees2/Dropbox/personal/matt/github/phd-thesis/index/bib/thesis.bib
header-includes:
    - \usepackage{setspace}\doublespacing
    - \usepackage{float}
    - \floatplacement{figure}{H}
    - \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}} \setcounter{section}{0} \renewcommand{\thesection}{S\arabic{section}}}
    - \usepackage{lineno}
linenumbers: false
csl: journal-of-applied-ecology.csl
geometry: margin=1.1in
fontsize: 12pt
output:
  bookdown::pdf_book:
    base_format: rticles::elsevier_article
editor_options: 
  chunk_output_type: inline
---

@@

## Study area and camera-trapping 
We compiled data from multiple smaller-scale camera-trap studies across two regions in south-west Victoria, Australia: the Glenelg region and Otway Ranges (Fig. \@ref(fig:diel-map)). Introduced foxes and cats are the only medium-large functional mammalian terrestrial carnivores here: native dingoes *Canis familiaris* are long-absent throughout, while tiger quolls *Dasyurus maculatus* are long-absent in the Glenelg region and likely functionally extinct in the Otway Ranges (last confirmed sighting in 2014). In broad sections of each region, government land managers conduct ongoing targeted lethal fox control for biodiversity conservation. Poison-baits containing 3 mg of sodium fluroacetate ('1080') are buried at a depth of 12 - 15 cm at 1-km intervals along accessible forest tracks and roads. Different road densities result in variable densities of poison-baits. Managers also frequently implement prescribed fire across both regions, primarily to reduce fuel loads to prevent large wildfires.

### Glenelg region
In the Glenelg region, large patches of natural vegetation are fragmented by pastoral farming and residential properties (Fig. \@ref(fig:diel-map)). Foxes in three distinct forest blocks in this region have been subject to poison-baiting since October 2005, with fortnightly bait replacements [@robley2014long]. These forest blocks, along with three similar, unbaited forest blocks to the north have been simultaneously surveyed annually under the 'Glenelg Ark' fox control program since 2005 [40 sites per block; @robley2020glenelg]. Hair-tubes were used to monitor species from 2005 - 2013 [presented in @robley2014long], replaced by camera-traps from 2013; here we present camera-trap data from 2013 - 2019 [@robley2020glenelg]. We also included a further 425 camera-trap deployments at unique locations from early 2018 (M.W.R. PhD surveys). This totals 2039 camera-trap deployments in the Glenelg region, collected in a control-impact experimental design (foxes had been continuously controlled for at 8 - 14 years in the treatment landscapes at the time of these surveys).

@@

### Camera-trap set-ups
All camera-trap deployments consisted of a Reconyx (Holmen, Wisconsin) brand camera-trap (white or infrared flash), attached to a tree or a metal picket, facing a lure. The Glenelg Ark and Otway Ark fox monitoring programs positioned camera-traps at least 40 cm above ground on a tree or a metal picket and angled downwards toward a lure approximately 1 - 1.5 m away [@robley2020glenelg; @robley2019otway]. The lures consisted of peanut butter, golden syrup and rolled oats mixed into a small ball, placed within a tea strainer or PVC pipe container and secured either to the ground, or 20 - 60 cm above ground on a wooden stake. The M.W.R. PhD surveys across both regions positioned camera-traps lower on a tree (around 15 - 30 cm above the ground) angled only slightly downwards toward a tuna oil lure approximately 2 - 2.5 m away [detailed in @rees2019unexpectedly]. Camera-traps were active for an average of 47 days (maximum 93 days), totalling 172,052 trap-nights.

@@

### How does predator overall activity and diel activity patterns vary across vegetation types? (model 2)
Predator activity varied across space; we hypothesised that this was partly due to differences in vegetation type, based both on the observed spatial patterns and because vegetation type is a major driver of understorey habitat structure and prey occurrence in these regions [@swan2015predicting; @hradsky2017bayesian]. To test whether the diel activity pattern of each predator varied among vegetation types, we identified the Ecological Vegetation Class group [hereafter 'vegetation type'; standard units for vegetation classification in Victoria; @delwp2020bioregions] for each unique camera-trap site, totalling eight vegetation types. As rainforests are interspersed (primarily in low lying gullies) at fine-scales throughout wet and damp forests in the south-eastern Otway Ranges, we merged them together (hereafter referred to as 'wet forests'). We then estimated predator activity across vegetation types using a hierarchical model specification: a global smoother for hour (i.e., average response) and group-level smoothers with shared wiggliness for the seven vegetation types ['model GS' detailed in @pedersen2019hierarchical]. We also included a random effect to account for differences in overall activity levels between the two regions.

@@

RESULTS
============
Overall, we collated 5449 and 2202 independent detections of foxes and cats, respectively (separated by at least 30 minutes) from 172,052 camera-trap nights (Table \@ref(tab:diel-tab1)).

### How does predator overall activity and diel activity patterns vary across space? (model 1)
Predator activity varied considerably across space and throughout the 24-hour daily cycle, and there was some variation in the predator diel activity patterns across space. On average, both predators showed similar diel activity patterns: mostly nocturnal with peaks in activity around sunrise and sunset (i.e., crepuscular; Fig. \@ref(fig:diel-veg)i). The main difference between the species was that fox activity peaked just after sunset and they were less likely to be active during the day than cats. Cats also tended to be more active at sunset relative to sunrise.

@@

Variation in diel activity patterns across space, as well as differences in overall activity between the predators, was strongest in the Otway Ranges. For example, overall fox activity (Fig. \@ref(fig:diel-space-o-fox)) and diel activity pattern strength (Fig. \@ref(fig:diel-space)b) were lowest in the south-west Otway Ranges, while feral cat overall activity (Fig. \@ref(fig:diel-space-o-cat)) and diel activity pattern strength (Fig. \@ref(fig:diel-space)b) were highest in that subregion.

### How does predator overall activity and diel activity patterns vary across vegetation types? (model 2)

@@

Diel activity patterns for foxes were similar across all vegetation type except wet forests; in wet forests, foxes were consistently active throughout the daily cycle (Fig. \@ref(fig:diel-veg)a). On the other hand, cats were nocturnal (and most active) in wet forests, but largely crepuscular in all other vegetation types (Fig. \@ref(fig:diel-veg)b). For both predators, the random effect for region (Glenelg or Otways) in the vegetation models shrank to near-zero, indicating all variation between the regions was explained by the vegetation covariate and site random intercept.

### Do feral cats avoid foxes in space or time? (model 3)
Cat spatial activity was relatively unaffected by the fox activity in both habitat types of the Otway Ranges; and, if anything, increased with increasing adjusted fox counts in the Glenelg region (Fig. \@ref(fig:diel-cat-fox)), indicating cats did not avoid foxes spatially.

@@

A key question in ecological theory is whether animals are evolutionary hardwired to occupy particular temporal niches, or have circadian rhythms that are responsive to changing environmental conditions and interactions with other species [@daan1981adaptive; @schoener1974compression; @lima1990behavioral]. Here we demonstrate that diel activity patterns are not fixed, but vary across space based on landscape context and fear. In our study, sympatric invasive predators had similar diel activity patterns when averaged across broad regions [i.e., high circular overlap; Fig. \@ref(fig:diel-space-time-marginal); as did @roshier2021space], but behaviours varied considerably within landscapes. Fox daily activity patterns were most strongly tied to the daily cycle in dry habitat types, but showed little diel activity pattern in the wet forests (Fig. \@ref(fig:diel-veg)a). In contrast, cats were mostly nocturnal in wet forests but crepuscular in dry vegetation types (Fig. \@ref(fig:diel-veg)b). Within broad habitat types, cats altered their diel activity patterns at sites with higher fox activity (Fig. \@ref(fig:diel-cat-fox)). Control programs that reduce invasive fox activity are therefore likely to change cat diel activity patterns, which in turn may alter cat impacts on native prey species. Quantifying changes in diel activity patterns provides important context for understanding species interactions, which is key for effective ecosystem management [@gaynor2021applied].

Shifting diel activity patterns may facilitate spatial coexistence of dominant and subordinate species [@carothers1984time]. For cats, altering diel activity patterns to less-preferred times of day may be worthwhile to persist in high-quality habitat. Few studies have demonstrated predator-induced shifts in diel activity [@kronfeld2003partitioning], but notably, ship rats *Rattus norvegicus* were also found to switch from nocturnal to diurnal behaviour in response to fox activity [@fenn1995use] and a similar nocturnal-diurnal shift was observed in American mink *Neovison vison* following the recolonisation of native predators [@harrington2009impact]. For cats in our study, a switch to diurnal behaviour where fox activity was high in the dry vegetation types may have been facilitated by the higher abundance of reptiles in these habitat types relative to wet forests [which are mostly diurnal; @woinarski2018reptiles]. At wet forest sites with high fox activity, cats concentrated their activity away from sunrise and sunset towards midnight, despite a diurnal shift appearing to similarly reduce the risk of a fox encounter. In this situation, we expect becoming more nocturnal to be favourable over a diurnal shift because this is when small mammals are active and cats would be least visible to foxes (cats here mostly had black or grey coats). Overall fox activity was twice as low in the wet forests relative to dry habitat types, and so cats were likely under less pressure to radically alter their diel activity patterns. Understanding how these potential avoidance behaviours impacts native prey is a key research priority to improve invasive predator management.

Cats may have avoided foxes in time, but we saw no sign of spatial avoidance (i.e. no fine-scale negative association between fox and cat overall activity (Fig. \@ref(fig:diel-cat-fox)b:c). Overlap in spatial activity at fine scales (i.e., at the level of camera-trap sites) may have been facilitated by temporal avoidance [@kronfeld2003partitioning]. However, we considered spatial avoidance across the entire survey duration (averaging 47 days). Cats may indeed avoid foxes spatially, but transiently on the scale of hours to days—after all, how is a cat to know where to avoid a fox without encountering signs of one? Short-term spatial avoidance is quite plausible given foxes mark territories using scats and odours, which cats could tangibly associate with high risk shortly after. Temporary spatial avoidance could be tested using decay curves [e.g., @niedballa2019assessing]. Alternatively, no sign of spatial avoidance between these invasive predators could also be an artefact of the quality of camera set-up and hence detectability. The 3667 camera-traps were deployed by numerous people, and the quality of set-ups differed considerably in terms of detecting predators. Camera-traps that angled only slightly downwards, rather than upwards or strongly downwards, seemed most effective at detecting both predator species (M.W Rees, personal observation). Predator interactions are routinely inferred through spatial associations between species; however such analyses are subject to numerous pitfalls which can make inference unreliable [reviewed in @guillaume2020co].

@@

Animal diel activity patterns can be complex, varying across space, habitat types and threat-levels. Despite telling an important story about how animals interact with each other and the environment, detection times are commonly discarded from statistical analyses of camera-trap data. In the rare instances that they are considered, diel activity patterns are predominantly estimated at the population-level, overlooking finer-scale behaviours that can affect fitness, survival and ecosystem-impacts. Our results demonstrate the importance of (a) considering diel activity in regards to species interactions, (b) modelling changes in animal behaviour rather than overlap with other species, and (c) testing avoidance behaviours within a joint spatiotemporal framework. Our study adds to the limited body of evidence that top predators can produce a landscape of fear which is powerful enough to reverse the diel activity patterns of subordinate species [@kronfeld2003partitioning].

@@

I had field assistance from: Shauni Omond, Shayne Neal, Asitha Samarawickrama, Shelley Thompson, Erin Harris, Hannah Killian, Lani Watson, Mark Dorman, Jack Davis, Carl Roffey, Bruce Edley, Larissa Oliveira Gonçalves, Ben Lake, Chantelle Geissler, Aviya Naccarella, Emily Gronow, Harley England, David Pitts, Annie Hobby, Louise Falls, Thomas McKinnon, Jimmy Downie, Marney Hradsky, Stephanie Samson, Robin Sinclair, Asmaa Alhusainan, Kelly Forrester, Tammana Wadawani, Emily Gregg, Hannah Edwards, Adam Beck, Vishnu Memnon, Sandy Lu, Pia Lentini, Nick Golding, Emily McColl-Gausden, Nina Page, Maggie Campbell-Jones, Kyle Quinn and Jack Dickson.

Surveys were conducted under University of Melbourne Animal Ethics Committee approval 1714119 and Victorian Government Department of Environment, Land Water and Planning Research Permit 10008273.

@@

# load records
fits_summaries <- read_csv("../derived_data/model_summaries.csv") %>%
  dplyr::select(-c(...1))
